---
title: "Final Plots"
author: "Michal Benjamin"
date: "2024-03-12"
output: html_document
---
#Bins per SEason GAD7
```{r setup, include=FALSE}
library(ggplot2)

# Convert "gad7.startDate" to Date format if it's not already in Date format
gad_data$gad7.startDate <- as.Date(gad_data$gad7.startDate)

# Function to get the season from a date
get_season <- function(date) {
  month <- as.integer(format(date, "%m"))
  season <- switch(month,
                   "12" = "Winter", "01" = "Winter", "02" = "Winter",
                   "03" = "Spring", "04" = "Spring", "05" = "Spring",
                   "06" = "Summer", "07" = "Summer", "08" = "Summer",
                   "09" = "Autumn", "10" = "Autumn", "11" = "Autumn",
                   "Unknown")
  return(season)
}

# Create a new column for the seasons
gad_data$season <- sapply(gad_data$gad7.startDate, get_season)

# Plot a discrete estimate graph measuring season vs GAD-7 Sum Score
plot <- ggplot(gad_data, aes(x = season, y = gad7.sum_score)) +
          stat_summary(fun = mean, geom = "point", size = 3, color = "skyblue") +
          stat_summary(fun.data = "mean_se", geom = "errorbar", width = 0.2) +
          labs(x = "Season", y = "Mean GAD-7 Sum Score", title = "Mean GAD-7 Sum Score by Season") +
          theme_minimal()

# Save the plot
ggsave("/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/gad7_season.png", plot)
library(dplyr)

# Assuming gad_data is your data frame with the season column
# Create season bins
gad_data <- gad_data %>%
  mutate(season_bins = case_when(
    season == "Winter" ~ "Winter",
    season == "Spring" ~ "Spring",
    season == "Summer" ~ "Summer",
    season == "Autumn" ~ "Autumn",
    TRUE ~ "Unknown" # Include the Unknown category
  ))

# Calculate the number of entries and unique entries per bin
entries_per_bin <- gad_data %>%
  group_by(season_bins) %>%
  summarise(entries = n(), unique_entries = n_distinct(ID))

# Print the result
print(entries_per_bin)

#Violin plot
library(ggplot2)
library(dplyr)

# Assuming gad_data is your data frame with the season column
# Create season bins
gad_data <- gad_data %>%
  mutate(season_bins = case_when(
    season == "Winter" ~ "Winter",
    season == "Spring" ~ "Spring",
    season == "Summer" ~ "Summer",
    season == "Autumn" ~ "Autumn",
    TRUE ~ "Unknown" # Include the Unknown category
  ))

# Plot the violin plot with average points
plot <- ggplot(gad_data, aes(x = season_bins, y = gad7.sum_score)) +
  geom_violin(trim = FALSE, fill = "skyblue") +
  stat_summary(fun = mean, geom = "point", color = "black", size = 2, shape = 18, fill = "black") + # Use diamond shape
  labs(x = "Season", y = "GAD-7 Sum Score", title = "Distribution of GAD-7 Sum Score by Season") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "white"))

# Save the plot
ggsave("/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/gad7_season_violin.png", plot)

# Calculate the number of entries and unique entries per bin
entries_per_bin <- gad_data %>%
  group_by(season_bins) %>%
  summarise(entries = n(), unique_entries = n_distinct(ID))

# Print the result
print(entries_per_bin)



#Raincloud plot
library(ggplot2)
library(dplyr)
library(ggdist)

# Assuming gad_data is your data frame with the season column
# Create season bins
gad_data <- gad_data %>%
  mutate(season_bins = case_when(
    season == "Winter" ~ "Winter",
    season == "Spring" ~ "Spring",
    season == "Summer" ~ "Summer",
    season == "Autumn" ~ "Autumn",
    TRUE ~ "Unknown" # Include the Unknown category
  ))

# Plot the violin plot with average points
plot <- ggplot(gad_data, aes(season_bins, gad7.sum_score)) +
  ggdist::stat_halfeye(adjust = .5, width = .3, .width = c(0.5, 1)) + 
  ggdist::stat_dots(side = "left", dotsize = .4, justification = 1.05, binwidth = .1)
  labs(x = "Season", y = "GAD-7 Sum Score", title = "Distribution of GAD-7 Sum Score by Season") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "white"))

# Save the plot
ggsave("/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/gad7_season_violin.png", plot)

# Calculate the number of entries and unique entries per bin
entries_per_bin <- gad_data %>%
  group_by(season_bins) %>%
  summarise(entries = n(), unique_entries = n_distinct(ID))

# Print the result
print(entries_per_bin)

#Raincloud plot
library(ggplot2)
library(dplyr)
library(ggdist)

# Assuming gad_data is your data frame with the season column
# Create season bins
gad_data <- gad_data %>%
  mutate(season_bins = case_when(
    season == "Winter" ~ "Winter",
    season == "Spring" ~ "Spring",
    season == "Summer" ~ "Summer",
    season == "Autumn" ~ "Autumn",
    TRUE ~ "Unknown" # Include the Unknown category
  ))

# Calculate average GAD-7 sum score per season
average_scores <- gad_data %>%
  group_by(season_bins) %>%
  summarise(avg_score = mean(gad7.sum_score))

# Plot the rain cloud plot with half-violin shape and dots representing averages
plot <- ggplot(gad_data, aes(season_bins, gad7.sum_score)) +
  geom_point(data = average_scores, aes(y = avg_score), size = 4, color = "black") + # Use average scores for points
  ggdist::stat_halfeye(adjust = .5, width = .3, .width = c(0.5, 1)) +
  labs(x = "Season", y = "GAD-7 Sum Score", title = "Distribution of Average GAD-7 Sum Score by Season") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "white"))

# Save the plot
ggsave("/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/gad7_season_raincloud.png", plot)

# Calculate the number of entries and unique entries per bin
entries_per_bin <- gad_data %>%
  group_by(season_bins) %>%
  summarise(entries = n(), unique_entries = n_distinct(ID))

# Print the result
print(entries_per_bin)



```

#Bins per SEason PHQ9
```{r setup, include=FALSE}
# Convert "phq9.startDate" to Date format if it's not already in Date format
phq_data$phq9.startDate <- as.Date(phq_data$phq9.startDate)

# Function to get the season from a date
get_season <- function(date) {
  month <- as.integer(format(date, "%m"))
  season <- switch(month,
                   "12" = "Winter", "01" = "Winter", "02" = "Winter",
                   "03" = "Spring", "04" = "Spring", "05" = "Spring",
                   "06" = "Summer", "07" = "Summer", "08" = "Summer",
                   "09" = "Autumn", "10" = "Autumn", "11" = "Autumn",
                   "Unknown")
  return(season)
}

# Create a new column for the seasons
phq_data$season <- sapply(phq_data$phq9.startDate, get_season)

# Plot a discrete estimate graph measuring season vs PHQ-9 Sum Score
plot <- ggplot(phq_data, aes(x = season, y = phq9.sum_score)) +
          stat_summary(fun = mean, geom = "point", size = 3, color = "salmon") +
          stat_summary(fun.data = "mean_se", geom = "errorbar", width = 0.2) +
          labs(x = "Season", y = "Mean PHQ-9 Sum Score", title = "Mean PHQ-9 Sum Score by Season") +
          theme_minimal()

# Save the plot
ggsave("/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/phq9_season.png", plot)

# Assuming phq_data is your data frame with the season column
# Create season bins
phq_data <- phq_data %>%
  mutate(season_bins = case_when(
    season == "Winter" ~ "Winter",
    season == "Spring" ~ "Spring",
    season == "Summer" ~ "Summer",
    season == "Autumn" ~ "Autumn",
    TRUE ~ "Unknown" # Include the Unknown category
  ))

# Calculate the number of entries and unique entries per bin
entries_per_bin <- phq_data %>%
  group_by(season_bins) %>%
  summarise(entries = n(), unique_entries = n_distinct(ID))

# Print the result
print(entries_per_bin)


```

#Bins per Month 
```{r setup, include=FALSE}
#Bins per SEason PHQ9
```{r setup, include=FALSE}
# Convert "phq9.startDate" to Date format if it's not already in Date format
phq_data$phq9.startDate <- as.Date(phq_data$phq9.startDate)

# Function to get the season from a date
get_season <- function(date) {
  month <- as.integer(format(date, "%m"))
  season <- switch(month,
                   "12" = "Winter", "01" = "Winter", "02" = "Winter",
                   "03" = "Spring", "04" = "Spring", "05" = "Spring",
                   "06" = "Summer", "07" = "Summer", "08" = "Summer",
                   "09" = "Autumn", "10" = "Autumn", "11" = "Autumn",
                   "Unknown")
  return(season)
}

# Create a new column for the seasons
phq_data$season <- sapply(phq_data$phq9.startDate, get_season)

# Plot a discrete estimate graph measuring season vs PHQ-9 Sum Score
plot <- ggplot(phq_data, aes(x = season, y = phq9.sum_score)) +
          stat_summary(fun = mean, geom = "point", size = 3, color = "salmon") +
          stat_summary(fun.data = "mean_se", geom = "errorbar", width = 0.2) +
          labs(x = "Season", y = "Mean PHQ-9 Sum Score", title = "Mean PHQ-9 Sum Score by Season") +
          theme_minimal()

# Save the plot
ggsave("/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/phq9_season.png", plot)

# Assuming phq_data is your data frame with the season column
# Create season bins
phq_data <- phq_data %>%
  mutate(season_bins = case_when(
    season == "Winter" ~ "Winter",
    season == "Spring" ~ "Spring",
    season == "Summer" ~ "Summer",
    season == "Autumn" ~ "Autumn",
    TRUE ~ "Unknown" # Include the Unknown category
  ))

# Calculate the number of entries and unique entries per bin
entries_per_bin <- phq_data %>%
  group_by(season_bins) %>%
  summarise(entries = n(), unique_entries = n_distinct(ID))

# Print the result
print(entries_per_bin)


```
```


## Bins per fortnight

You can also embed plots, for example:

```{r pressure, echo=FALSE}
# Load required libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)

# Define a function to calculate two-week periods starting from the beginning of the year
two_week_period <- function(date) {
  period <- floor((as.numeric(format(date, "%j")) - 1) / 14) + 1
  return(period)
}

# Read in the PHQ data
phq_data <- readRDS("/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/phq_clinical_weather.rds")

# Filter out rows with NA values in the phq9.startDate and phq9.sum_score columns
filtered_phq_data <- phq_data %>%
  filter(!is.na(phq9.startDate), !is.na(phq9.sum_score))

# Extract month and day from phq9.startDate column, ignoring the year section
filtered_phq_data$start_date <- as.Date(filtered_phq_data$phq9.startDate)
filtered_phq_data$month_day <- format(filtered_phq_data$start_date, "%m-%d")

# Add a two-week period column based on the month-day combinations
filtered_phq_data$two_week_period <- two_week_period(as.Date(filtered_phq_data$month_day, "%m-%d"))

# Calculate the average sum score for each two-week period in PHQ data
average_phq_scores <- filtered_phq_data %>%
  group_by(two_week_period) %>%
  summarise(avg_sum_score_phq = mean(phq9.sum_score, na.rm = TRUE))

# Read in the GAD data
gad_data <- readRDS("/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/gad_clinical_weather.rds")

# Filter out rows with NA values in the gad7.startDate and gad7.sum_score columns
filtered_gad_data <- gad_data %>%
  filter(!is.na(gad7.startDate), !is.na(gad7.sum_score))

# Extract month and day from gad7.startDate column, ignoring the year section
filtered_gad_data$start_date <- as.Date(filtered_gad_data$gad7.startDate)
filtered_gad_data$month_day <- format(filtered_gad_data$start_date, "%m-%d")

# Add a two-week period column based on the month-day combinations
filtered_gad_data$two_week_period <- two_week_period(as.Date(filtered_gad_data$month_day, "%m-%d"))

# Calculate the average sum score for each two-week period in GAD data
average_gad_scores <- filtered_gad_data %>%
  group_by(two_week_period) %>%
  summarise(avg_sum_score_gad = mean(gad7.sum_score, na.rm = TRUE))

# Define the file path for saving the plot
file_path <- "/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/gad7_phq9_average_sum_scores.png"

# Plot the graph with separate lines for GAD-7 and PHQ-9
plot <- ggplot() +
  geom_line(data = average_gad_scores, aes(x = two_week_period, y = avg_sum_score_gad, color = "GAD-7")) +
  geom_line(data = average_phq_scores, aes(x = two_week_period, y = avg_sum_score_phq, color = "PHQ-9")) +
  scale_color_manual(values = c("GAD-7" = "skyblue", "PHQ-9" = "salmon")) +  # Set colors
  labs(title = "Average GAD-7 and PHQ-9 Sum Scores Over Time",
       x = "Two-Week Period",
       y = "Average Sum Score",
       color = "Scale") +
  theme_minimal()

# Save the plot
ggsave(file_path, plot, width = 10, height = 6, units = "in")
# Plot the graph for GAD-7
gad_plot <- ggplot(average_gad_scores, aes(x = as.factor(two_week_period), y = avg_sum_score_gad)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Average GAD-7 Sum Scores Over Time",
       x = "Two-Week Period",
       y = "Average Sum Score") +
  theme_minimal()

# Plot the graph for PHQ-9
phq_plot <- ggplot(average_phq_scores, aes(x = as.factor(two_week_period), y = avg_sum_score_phq)) +
  geom_bar(stat = "identity", fill = "salmon") +
  labs(title = "Average PHQ-9 Sum Scores Over Time",
       x = "Two-Week Period",
       y = "Average Sum Score") +
  theme_minimal()

# Define file paths
gad_file_path <- "/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/gad_fortnight.png"
phq_file_path <- "/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/phq_fortnight.png"

# Save the plots
ggsave(gad_file_path, gad_plot, width = 10, height = 6, units = "in")
ggsave(phq_file_path, phq_plot, width = 10, height = 6, units = "in")

```

#Bins per Daylight
```{r Clear global environment}
remove(list = ls())

#Bins per daylight
# Create daylight bins with 2-hour intervals
phq_data <- phq_data %>%
  mutate(daylight_bins = cut(phq9_average_daylight_hr, breaks = c(seq(7, 17, by = 2), 24), include.lowest = TRUE))

# Reorder the levels of daylight_bins in ascending numerical order
phq_data$daylight_bins <- factor(phq_data$daylight_bins, levels = unique(phq_data$daylight_bins))

# Shorten the names of the daylight bins
levels(phq_data$daylight_bins) <- c("(7,9]", "(9,11]", "(11,13]", "(13,15]", "(15,17]", "17+")

# Calculate the number of entries and unique entries per bin
entries_per_bin_daylight <- phq_data %>%
  group_by(daylight_bins) %>%
  summarise(entries = n(), unique_entries = n_distinct(ID))

# Print the result
print(entries_per_bin_daylight)

# Plot a discrete estimate graph measuring daylight vs PHQ-9 Sum Score with error bars
plot_daylight <- ggplot(phq_data, aes(x = daylight_bins, y = phq9.sum_score)) +
  stat_summary(fun = mean, geom = "point", size = 3, color = "salmon") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = 0.2) +
  labs(x = "Daylight Hours", y = "Mean PHQ-9 Sum Score", title = "Mean PHQ-9 Sum Score by Daylight Hours") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# Save the plot
ggsave("/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/phq9_daylight.png", plot_daylight)
```

#Sunshine hours
```{r Clear global environment}
# Load required libraries
library(dplyr)
library(ggplot2)

#Bins per sunlight
# Create sunlight bins with 1-hour intervals
gad_data <- gad_data %>%
  mutate(sun_bins = cut(gad7_average_sun_hr, breaks = c(1, 2, 3, 4, 5, 6, 7), include.lowest = TRUE))

# Reorder the levels of sun_bins in ascending numerical order
gad_data$sun_bins <- factor(gad_data$sun_bins, levels = c("(1,2]", "(2,3]", "(3,4]", "(4,5]", "(5,6]", "(6,7]"))

# Calculate the number of entries and unique entries per bin
entries_per_bin_sunlight <- gad_data %>%
  group_by(sun_bins) %>%
  summarise(entries = n(), unique_entries = n_distinct(ID))

# Print the result
print(entries_per_bin_sunlight)

# Plot a discrete estimate graph measuring sunlight vs GAD-7 Sum Score with error bars
plot_sunlight <- ggplot(gad_data, aes(x = sun_bins, y = gad7.sum_score)) +
  stat_summary(fun = mean, geom = "point", size = 3, color = "skyblue") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = 0.2) +
  labs(x = "Sunlight Hours", y = "Mean GAD-7 Sum Score", title = "Mean GAD-7 Sum Score by Sunlight Hours") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# Save the plot
ggsave("/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/gad7_sunlight.png", plot_sunlight)

```

#Plots per wave
```{r Clear global environment}
library(dplyr)
library(ggplot2)

# Filter data to include only waves 1 to 18
phq_data_filtered <- phq_data %>%
  filter(!is.na(wave_assigned) & wave_assigned != "wave19" & wave_assigned != "wave20" & wave_assigned != "wave21") 

# Convert the wave_assigned column to numeric for correct sorting
phq_data_filtered$wave_assigned <- gsub("wave", "", phq_data_filtered$wave_assigned)
phq_data_filtered$wave_assigned <- as.numeric(phq_data_filtered$wave_assigned)

# Arrange the data by wave in ascending order
phq_data_filtered <- phq_data_filtered %>%
  arrange(wave_assigned)

# Calculate the number of entries and unique entries per wave
entries_per_wave <- phq_data_filtered %>%
  group_by(wave_assigned) %>%
  summarise(entries = n(), unique_entries = n_distinct(ID))

# Print the result
print(entries_per_wave)

# Calculate the mean sumscore per wave with error bars
sumscore_per_wave <- phq_data_filtered %>%
  group_by(wave_assigned) %>%
  summarise(mean_sumscore = mean(phq9.sum_score, na.rm = TRUE),
            sd_sumscore = sd(phq9.sum_score, na.rm = TRUE),
            se_sumscore = sd_sumscore / sqrt(n()))

# Plot the mean sumscore per wave with error bars
plot_wave_sumscore <- ggplot(sumscore_per_wave, aes(x = factor(wave_assigned), y = mean_sumscore)) +
  geom_point(size = 3, color = "salmon") +
  geom_errorbar(aes(ymin = mean_sumscore - se_sumscore, ymax = mean_sumscore + se_sumscore), 
                width = 0.2, color = "salmon") +
  labs(x = "Wave", y = "Mean PHQ-9 Sum Score", title = "Mean PHQ-9 Sum Score per Wave") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Save the plot
ggsave("/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/phq9_wave_sumscore.png", plot_wave_sumscore)

```

#Regression winter vs lockdown
```{r Clear global environment}
# Assuming you have already loaded your dataset and imported necessary libraries
# First, create a new variable indicating whether the wave occurred during lockdown or winter
phq_data$lockdown <- ifelse(phq_data$wave_assigned %in% c("wave1", "wave12", "wave13"), "lockdown", "non-lockdown")
phq_data$winter <- ifelse(phq_data$wave_assigned %in% c("wave10", "wave11", "wave12", "wave13"), "winter", "non-winter")

# Then, calculate the average sum scores for each wave
lockdown_avg <- aggregate(phq9.sum_score ~ lockdown, data = phq_data, FUN = mean)
winter_avg <- aggregate(phq9.sum_score ~ winter, data = phq_data, FUN = mean)

# Now, let's run a regression analysis
model <- lm(phq9.sum_score_baseline ~ lockdown + winter, data = phq_data)

# Print summary of the regression model
summary(model)

# Assuming you have already loaded your dataset and imported necessary libraries
# First, create a new variable indicating whether the wave occurred during lockdown or winter
gad_data$lockdown <- ifelse(gad_data$wave_assigned %in% c("wave1", "wave12", "wave13"), "lockdown", "non-lockdown")
gad_data$winter <- ifelse(gad_data$wave_assigned %in% c("wave10", "wave11", "wave12", "wave13"), "winter", "non-winter")

# Then, calculate the average sum scores for each wave
lockdown_avg <- aggregate(gad_data$gad7.sum_score ~ gad_data$lockdown, data = gad_data, FUN = mean)
winter_avg <- aggregate(gad_data$gad7.sum_score ~ gad_data$winter, data = gad_data, FUN = mean)

# Now, let's run a regression analysis
model <- lm(gad7.sum_score ~ lockdown + winter, data = gad_data)

# Print summary of the regression model
summary(model)




#########
# Load required libraries
library(dplyr)

# Assuming you have a dataset named gad_data with columns:
# gad7_sum_score: GAD-7 sum scores
# wave_assigned: Assigned wave for each observation

# Create indicator variables for winter waves 10-13, summer lockdown (wave 1), and winter lockdown waves 12-13 based on wave_assigned
gad_data <- gad_data %>%
  mutate(
    winter_waves = ifelse(wave_assigned %in% c(10, 11, 12, 13), 1, 0),
    summer_lockdown = ifelse(wave_assigned == 1, 1, 0),
    winter_lockdown = ifelse(wave_assigned %in% c(12, 13), 1, 0)
  )

# Perform regression analysis
regression_model <- lm(gad7_sum_score ~ winter_waves + summer_lockdown + winter_lockdown, data = gad_data)

# View summary of regression results
summary(regression_model)

# Optionally, you can test for overall significance of the model
# using an ANOVA test
anova(regression_model)

# Optionally, you can perform additional diagnostics such as
# checking for multicollinearity and homoscedasticity
# For example, to check for multicollinearity:
library(car)  # Load car package for VIF calculation
vif(regression_model)

```

Clear global environment prior to initiation
```{r Clear global environment}
remove(list = ls())
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
