---
title: "Regression analysis"
author: "Michal Benjamin"
date: "2024-03-29"
output: html_document
---
#Rezidualizing Data- GAD7
```{r setup, include=FALSE}
# Load necessary libraries
library(dplyr)

# Define residualization function
residualize_data <- function(data, outcome_var, predictors) {
  # Remove rows with missing values
  data <- na.omit(data[, c(outcome_var, predictors)])
  
  # Convert factors to character vectors
  data <- lapply(data, function(x) if(is.factor(x)) as.character(x) else x)
  data <- as.data.frame(data)
  
  # Fit a regression model
  model <- lm(data[[outcome_var]] ~ ., data = data[, c(outcome_var, predictors)])
  
  # Get residuals
  residuals <- residuals(model)
  
  # Add residuals to the original data
  data$residuals <- residuals
  
  # Return the residualized data
  return(data)
}

# Read in the data
gad_data <- readRDS("/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/gad_clinical_weather_saved.rds")

# Specify outcome variable and predictors
outcome_var <- "gad7.sum_score"
predictors <- c("age", "sex", "wave_assigned")  # Including "wave_assigned" as a predictor

# Perform residualization
residualized_gad_data <- residualize_data(gad_data, outcome_var, predictors)

# Rename the residuals column
colnames(residualized_gad_data)[ncol(residualized_gad_data)] <- paste(outcome_var, "residuals", sep = "_")

# Specify file path
file_path <- "/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/residualized_gad_data.rds"

# Save the residualized data as an RDS file
saveRDS(residualized_gad_data, file_path)


```

## Residualizing data- PHQ9

```{r cars}
# Load necessary libraries
library(dplyr)

# Define residualization function
residualize_data <- function(data, outcome_var, predictors) {
  # Remove rows with missing values
  data <- na.omit(data[, c(outcome_var, predictors)])
  
  # Convert factors to character vectors
  data <- lapply(data, function(x) if(is.factor(x)) as.character(x) else x)
  data <- as.data.frame(data)
  
  # Fit a regression model
  model <- lm(data[[outcome_var]] ~ ., data = data[, c(outcome_var, predictors)])
  
  # Get residuals
  residuals <- residuals(model)
  
  # Add residuals to the original data
  data$residuals <- residuals
  
  # Return the residualized data
  return(data)
}

# Read in the data
phq_data <- readRDS("/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/phq_clinical_weather_saved.rds")

# Specify outcome variable and predictors
outcome_var <- "phq9.sum_score"
predictors <- c("age", "sex")

# Perform residualization
residualized_phq_data <- residualize_data(phq_data, outcome_var, predictors)

# Rename the residuals column
colnames(residualized_phq_data)[ncol(residualized_phq_data)] <- paste(outcome_var, "residuals", sep = "_")

# Specify file path
file_path <- "/users/k21043154/project/seasonal_internalising_2024/data/phenotypic/clinical/best_data/residualized_phq_data.rds"

# Save the residualized data as an RDS file
saveRDS(residualized_phq_data, file_path)

```

## Regression Model



```{r pressure, echo=FALSE}
#Regression winter vs lockdown
```{r Clear global environment}
# Assuming you have already loaded your dataset and imported necessary libraries
# First, create a new variable indicating whether the wave occurred during lockdown or winter
residualized_phq_data$lockdown <- ifelse(phq_data$wave_assigned %in% c("wave1", "wave12", "wave13"), "lockdown", "non-lockdown")
residualized_phq_data$winter <- ifelse(phq_data$wave_assigned %in% c("wave10", "wave11", "wave12", "wave13"), "winter", "non-winter")

# Then, calculate the average sum scores for each wave
lockdown_avg <- aggregate(phq9.sum_score ~ lockdown, data = phq_data, FUN = mean)
winter_avg <- aggregate(phq9.sum_score ~ winter, data = phq_data, FUN = mean)

# Now, let's run a regression analysis
model <- lm(phq9.sum_score_baseline ~ lockdown + winter, data = phq_data)

# Print summary of the regression model
summary(model)

# Assuming you have already loaded your dataset and imported necessary libraries
# First, create a new variable indicating whether the wave occurred during lockdown or winter
residualized_gad_data$lockdown <- ifelse(gad_data$wave_assigned %in% c("wave1", "wave12", "wave13"), "lockdown", "non-lockdown")
residualized_gad_data$winter <- ifelse(gad_data$wave_assigned %in% c("wave10", "wave11", "wave12", "wave13"), "winter", "non-winter")

# Then, calculate the average sum scores for each wave
lockdown_avg <- aggregate(gad_data$gad7.sum_score ~ gad_data$lockdown, data = gad_data, FUN = mean)
winter_avg <- aggregate(gad_data$gad7.sum_score ~ gad_data$winter, data = gad_data, FUN = mean)

# Now, let's run a regression analysis
model <- lm(gad7.sum_score ~ lockdown + winter, data = gad_data)

# Print summary of the regression model
summary(model)




#########
# Load required libraries
library(dplyr)

# Assuming you have a dataset named gad_data with columns:
# gad7_sum_score: GAD-7 sum scores
# wave_assigned: Assigned wave for each observation

# Create indicator variables for winter waves 10-13, summer lockdown (wave 1), and winter lockdown waves 12-13 based on wave_assigned
gad_data <- gad_data %>%
  mutate(
    winter_waves = ifelse(wave_assigned %in% c(10, 11, 12, 13), 1, 0),
    summer_lockdown = ifelse(wave_assigned == 1, 1, 0),
    winter_lockdown = ifelse(wave_assigned %in% c(12, 13), 1, 0)
  )

# Perform regression analysis
regression_model <- lm(gad7_sum_score ~ winter_waves + summer_lockdown + winter_lockdown, data = gad_data)

# View summary of regression results
summary(regression_model)

# Optionally, you can test for overall significance of the model
# using an ANOVA test
anova(regression_model)

# Optionally, you can perform additional diagnostics such as
# checking for multicollinearity and homoscedasticity
# For example, to check for multicollinearity:
library(car)  # Load car package for VIF calculation
vif(regression_model)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
